---
title: "R Notebook"
output: html_notebook
---
1. Load in the libraries you need, and look at the backpack data. Take note of variable names and dimensions.
```{r}
library(tidyverse)
backpack <- CodeClanData::backpack
```

```{r}
names(backpack)
```
```{r}
dim(backpack)
```

2. What is the distribution of backpack weights that students are carrying? Choose an appropriate graph to show the distribution.

```{r}
ggplot(backpack) +
  geom_bar(aes(x = backpack_weight, fill = sex), color = "white") +
  labs(
    x = "Backpack Weight (kg)",
    y = "Count of Backpacks per Weight",
    title = "Weight of Backpacks Carried by Students",
    fill = "Sex"
  ) +
  scale_x_continuous(breaks = c(0, 5, 10, 15, 20, 25, 30, 35, 40)) +
  scale_y_continuous(breaks = c(0, 2, 4, 6, 8, 10, 12, 14)) +
  theme_light()
#The majority of students carry a backpack weighing between 5 and 15kg.
```
3. Are there differences in backpack weight between sex?

```{r}
ggplot(backpack) +
  aes(x = backpack_weight, fill = sex, colour = "white") +
  scale_fill_manual(values=c('orange', 'purple')) +
  geom_histogram(bins = 20, col = "white", show.legend = FALSE) + 
  theme_light() + 
  labs(y = "number of backpacks", x = "weight") + 
  facet_wrap(~ sex, nrow = 2)
#The data shows more females carrying bags than males, however there is a right skewed distribution for each sex. There are more males carrying bags over 25kg than females. 
```


4. What is the relationship between body weight and backpack weight?
```{r}
ggplot(backpack) +
  geom_point(aes(x = body_weight, y = backpack_weight), color = "steel blue") +
  labs(
    x = "Body Weight",
    y = "Backpack  Weight",
    title = "Weight of Backpacks Carried by Body Weight"
  ) 
#There is no clear relationship between backpack weight and student weight.
```
5. Is this relationship different between men and women? Choose an appropriate plot and interpret your findings.
```{r}
ggplot(backpack) +
  geom_point(aes(x = body_weight, y = backpack_weight, color = sex)) +
  labs(
    x = "Body Weight",
    y = "Backpack  Weight",
    title = "Weight of Backpacks Carried by Body Weight",
fill = "Sex") +
  scale_fill_manual(values=c('orange', 'purple')) +
  theme_light() +
  facet_wrap(~ sex)

#Females carry backpacks which are lighter than those carried by males
```

6. What proportion of students report back pain?
```{r}
backpack <- backpack %>%
  mutate(backpain_labelled = recode(back_problems, 
                       "0" = "No", 
                       "1" = "Yes"))
ggplot(backpack) +
  geom_bar(aes(x = backpain_labelled, fill = backpain_labelled)) +
   labs(
    x = "Back Problems",
    y = "Number of Students",
    title = "Proportion of Students with Back Problems",
fill = "Sex") +
  ylim(0,100) +
  scale_fill_manual(values=c('orange', 'purple')) +
  theme_light() +
  theme(legend.position = "none")
#One third of students are reporting back pain.
```



7. Is this proportion different between men and women?
```{r}
ggplot(backpack) +
  geom_bar(aes(x = backpain_labelled, fill = backpain_labelled)) +
   labs(
    x = "Back Problems",
    y = "Number of Students",
    title = "Proportion of Students with Back Problems",
fill = "Sex") +
  facet_wrap(~sex) +
  scale_fill_manual(values=c('orange', 'purple')) +
  theme_light() +
  theme(legend.position = "none")
#More females report back pain than males, possibly due to females carrying backpacks of a similar weight to males but with the disadvantage of lower body weights. 
```

8.
Now you’ve investigated your data separately, create plots to describe the relationship between backpack weight, body weight, gender and reported backpain.

Try both backpack_weight and body_weight as well as the ratio.

```{r}

pain_labels <- c("backpain: no", "backpain: yes")
names(pain_labels) <- c("No", "Yes")

sex_labels <- c("sex: female", "sex: male")
names(sex_labels) <- c("Female", "Male")


ggplot(backpack) + 
  aes(x = body_weight, y = backpack_weight, color = backpain_labelled) + 
  geom_point() + 
  facet_grid(sex ~ backpain_labelled, 
             labeller = labeller(backpain_labelled = pain_labels, sex = sex_labels)) + 
  labs(x = "\n body weight", y = "backpack weight \n") +
      scale_color_manual(values=c('#f0ca41', '#5022bd'))+
theme(legend.position = "none")

  
```
```{r}

ggplot(backpack) + 
  geom_histogram(aes(x = ratio, fill = backpain_labelled), bins = 20, col = "white") + 
  facet_grid(sex ~ backpain_labelled, 
             labeller = labeller(backpain_labelled = pain_labels, sex = sex_labels)) + 
  labs(x = "\n body weight: backpack weight ratio") +
theme(legend.position = "none") +
 scale_fill_brewer(palette = "Set2")
```

