---
title: "R Notebook"
output: html_notebook
---

# MVP 

1. Load the `code_clan_tweets.csv` data. Find the number of rows, columns, and list all the variable names.

```{r}
library(tidyverse)
library(janitor)
```

```{r}
#reading in the data
code_clan_tweets <- read_csv("data/code_clan_tweets.csv")

```

```{r}
#finding the number of rows and columns
dim(code_clan_tweets)
```

```{r}
#finding the number of columns
ncol(code_clan_tweets)
```

```{r}
#finding the number of rows
nrow((code_clan_tweets))
```

```{r}
#finding the variable names
names(code_clan_tweets)
```

```{r}
#having a look at the data types
glimpse(code_clan_tweets)
```

```{r}
#having a preview of the data (top 10 rows)

head(code_clan_tweets, 10)
```

```{r}
#having a preview of the data (bottom 10 rows)

tail(code_clan_tweets, 10)
```

2. Find the total number of favourited tweets (stored in `favorite_count`) that CodeClan tweets have got. Don't add a tweet's `favorite_count` to the total when the tweet was quoted (see the `is_quote` variable).   

```{r}
total_favourited_tweets <-
  code_clan_tweets %>% 
  filter(is_quote == FALSE) %>% 
  group_by(screen_name) %>% 
  summarise(total_favourited_tweets = sum(favorite_count)) 
```




3. Summarise the mean number of retweets (`retweet_count`), split by the type of platform that was used to tweet (`source`). Again, omit quoted tweets from the mean.

```{r}
mean_number_retweets <- 
  code_clan_tweets %>% 
  filter(is_quote == FALSE) %>% 
  group_by(source) %>% 
  summarise(mean_number_retweets = mean(retweet_count)) 
```


4. Count the total number of likes (i.e. total of `favorite_count`), split by `media` type, and arrange them from most likes to least. Assume that any tweets without a listed `media` type are of type "text".  

```{r}
total_number_likes <- 
  code_clan_tweets %>% 
  mutate(media_type2 = replace_na(media_type, "text")) %>% 
  group_by(media_type2) %>% 
  summarise(total_number_likes = sum(favorite_count)) %>% 
  arrange(desc(total_number_likes)) 

total_number_likes
```



5. Find the mean number of characters that a CodeClan tweet contains.  
```{r}
#find total number of characters for each CodeClan tweet
total_number_characters <- 
  code_clan_tweets %>% 
   filter(screen_name == "CodeClanScot")  %>% 
  group_by(text) %>% 
  summarise(total_characters = (str_count(text)))

total_number_characters

#find the mean number of characters that a CodeClan tweet contains
mean_number_characters <-
  total_number_characters %>% 

  summarise(mean_number_characters = mean(total_characters))

mean_number_characters

```


6. The `code_clan_info.csv` data file contains `status_url` along with other info. Load this in, and join it to the `code_clan_tweets` tibble, so that you have a `status_url` for each tweet. Decide which variable to join the tibbles on.  

```{r}
code_clan_info <-
read_csv("data/code_clan_info.csv")
```

```{r}
codeclan_join <- 
left_join(code_clan_info, code_clan_tweets, by = "tweet_id")  
  
  view(codeclan_join)
```



7. From your new joined data, create a new tibble `codeclan_hashtags` containing only the `tweet_id` and hashtags in lowercase for analysis. Keep only those tweets with hashtags.

```{r}

codeclan_hashtags <- codeclan_join %>% 
  select(tweet_id, hashtags) %>% 
   mutate(hashtags_lowcase = str_to_lower(hashtags)) %>% 
  select(-hashtags) %>% 
  drop_na(hashtags_lowcase) 

codeclan_hashtags

```


```{r}
codeclan_hashtags <-
  codeclan_join %>% 
 

codeclan_hashtags <- codeclan_join %>% 
  select(tweet_id, hashtags) %>% 
  drop_na() 

codeclan_hashtags
```



