---
title: "Distributions Homework"
output: html_notebook
---

```{r}
#Loading tidyverse to be able to read csv file
library(tidyverse)
#Loading janitor to be able to clean names
library(janitor)
```
1. 
```{r}
#Reading in the data and cleaning variable names
mobile_phones <- read_csv("data/20190928-items.csv") %>%  clean_names()

head(mobile_phones)

mobile_brands <- unique(mobile_phones$brand)

mobile_brands
```

2. Find the brand with the highest numbers of phones in the dataset.
```{r}
mobile_phones %>% 
  group_by(brand) %>% 
  summarise(number_of_phones = n()) %>% 
  arrange(desc(number_of_phones))
```

```{r}
mobile_phones %>% 
  ggplot(aes(x = brand)) +
  geom_bar() +
  theme_light() +
   theme(axis.text.x = element_text(angle = 55, hjust = 1)) +
  labs(
    title = "Number of Phones by Brand"
  ) +
  xlab("Brand") +
  ylab("Number of Phones") 
```
3. For your top brand, plot the distribution of phone ratings as a probability density, overlaying a fitted normal distribution. Do you think the normal distribution provides a good model of these ratings?

The graph is left skewed and the normal distribution does not provide a good model of the ratings.
```{r}
samsung_stats <- 
mobile_phones %>% 
  #Filtering results to show top brand only.
  filter(brand == "Samsung") %>% 
  summarise(
    num = n(),
    mean = mean(rating),
    sd = sd(rating)
  )

mobile_phones %>% 
  filter(brand == "Samsung") %>% 
  ggplot(aes(x = rating)) +
  geom_histogram(aes(y = ..density..), col = "white", bins = 25) +
  #creating line showing normal distribution
  stat_function(
    #normal distribution
    fun = dnorm,
    args = list(
      mean = samsung_stats$mean,
      sd = samsung_stats$sd
    ),
    col = "red"
  ) +
  theme_light() +
  ylab("Probability Density") +
  xlab("Phone Rating") +
  labs(
    title = "Distribution of Ratings for Samsung Phones",
    subtitle = "Probability Density Overlaying a Fitted Normal Distribution"
  ) 
```

4. We can be more quantitative in comparing the distribution of top brand ratings with a normal distribution. You calculated the mean() and sd() of your distribution in an earlier question. Use these values to calculate the proportions of ratings within one-, two- and three standard deviations of the mean. Compare these proportions with those you expect for a normal distribution.

i. Finding the proportion of ratings within one standard deviation of the mean

```{r}
samsung_ratings <- mobile_phones %>% 
  select(brand, rating) %>% 
   filter(brand == "Samsung")

mobile_phones %>%
  filter(brand == "Samsung") %>% 
  filter(rating >= samsung_stats$mean - samsung_stats$sd) %>%
  filter(rating <= samsung_stats$mean + samsung_stats$sd) %>%
  summarise(prop_within_1sd = n() / nrow(samsung_ratings))
```

ii. Within 2 standard deviations
```{r}
mobile_phones %>%
  filter(brand == "Samsung") %>% 
  filter(rating >= samsung_stats$mean - 2* samsung_stats$sd) %>%
  filter(rating <= samsung_stats$mean + 2* samsung_stats$sd) %>%
  summarise(prop_within_2sd = n() / nrow(samsung_ratings))
```
Within 3 standard deviations
```{r}
mobile_phones %>%
  filter(brand == "Samsung") %>% 
  filter(rating >= samsung_stats$mean - 3* samsung_stats$sd) %>%
  filter(rating <= samsung_stats$mean + 3* samsung_stats$sd) %>%
  summarise(prop_within_3sd = n() / nrow(samsung_ratings))
```

```{r}
#Creating scale column
samsung_scaled <- mobile_phones %>% 
  filter(brand == "Samsung")  %>% 
  mutate(z_rating = scale(rating))

samsung_scaled
```


```{r}
#Filtering out the results for 3 standard deviations away from the mean to find the outliers
samsung_scaled %>% 
  filter(!between(z_rating, left = -3, right = 3))
#There are 8 outliers.
```




For a normal distribution, expect 68%
 of values to fall within one standard deviation of the mean. Here, 79%
 of ratings are within this range, so our distribution is slightly too central.

Adding another standard deviation to either side should increase the proportion of values by 27%
 to 95%
 overall in a normal distribution. Our distribution adds 13%
, going to 92%
 overall, so there are too few values in the range from one- to two standard deviations.

Finally, in a normal distribution, adding a further standard deviation to either side should increase the proportion of values by 4.7%
 to 99.7%
 overall. Our distribution adds 6%
 to 98%, so this is pretty close.
 
 2. Extension
```{r}
qqnorm(samsung_ratings$rating)
qqline(samsung_ratings$rating)

```
 
```{r}

qqPlot(samsung_ratings$rating)
```
 
 