---
title: "Linear Regression"
output:
  html_document:
    df_print: paged
---

```{r}
library(tidyverse)
library(janitor)
library(modelr)
project_management <- read.csv("data/project_management.csv")
```

Plotting actual_length as the dependent variable (y by convention), and estimated_length as the independent variable (x by convention).
```{r}
project_management %>% 
  ggplot(aes(x = estimated_length, y = actual_length)) +
  geom_point()
```

```{r}
#Calculating the correlation coefficient of estimated_length and actual_length and interpreting the obtained value.
project_management %>% 
  summarise(cor(estimated_length, actual_length))

#The correlation of the coefficient of estimated_length and actual_length is 0.8046 which means the correlation is very strong.
```

Creating a simple linear regression using actual_length as the dependent variable, and estimated_length as the independent variable. 

```{r}
model <- lm(formula = actual_length ~ estimated_length, data = project_management)

model

summary(model)
```
The equation of the regression line is:
actual_length = 1.223 * estimated_length + 1.416

A 1 day increase in estimated_length is associated with a 1.2235 day increase in actual_length, i.e. the company is underestimating job lengths.

The r^2 (coefficient of determination) value (0.6475) means that 64.75% of the variation in actual_length can be predicted from (explained by) the variation in estimated_length.
```{r}
library(broom)
library(ggfortify)

project_management %>% 
  ggplot(aes(x = estimated_length, y = actual_length)) +
  geom_point() +
  geom_abline(
    intercept = model$coefficients[1],
    slope = model$coefficients[2],
    col = "red"
  )

#or 
#project_management %>% 
#ggplot(aes(x = estimated_length)) +
#  geom_point(aes(y = actual_length)) +
#  geom_line(aes(y = pred), col = "red")

```

```{r}
glance(model)
```

The relationship is statistically significant as the p-value of the slope is much less than the conventional value of α = 0.05.

```{r}
tidy(model)
```

```{r}
autoplot(model)
```
Aside from outliers (certainly point 5, and perhaps point 18), the diagnostic plots looks reasonable. The ‘Residuals vs Fitted’ plot reveals independent residuals, the ‘Normal Q-Q’ plot looks fine, and there is no systematic upward or downward trend in the the ‘Scale-Location’ plot.

```{r}
project_management %>% 
  ggplot(aes(x = estimated_length, y = actual_length)) +
  geom_point() +
  geom_abline(
    intercept = model$coefficients[1],
    slope = model$coefficients[2],
    col = "red"
  ) + 
  geom_text(aes(label = 1:nrow(project_management)), nudge_x = 0.2, nudge_y = 1)
```
Interpreting the Residuals vs Leverage diagnostic plot.
Points 5 & 18 are outliers.

```{r}
#Plot of Cook’s distance
autoplot(model, which = 4)

#The Cook's distance plot shows that point 5 is an influential outlier as it lies above 1. Point 18 is a non-influential outlier.
```


```{r}
#Creating a two-by-two diplay of the model plots (using Base R function)
par(mfrow = c(2,2))
plot(model)
```

Obtaining the intercept and regression coefficient of estimated_length for a simple linear model fitted to data omitting one of the non-influential outlier points.
```{r}
non_influential_omitted <- project_management %>% 
  slice(-18)

non_influential_omitted_model <- lm(formula = actual_length ~ estimated_length, data = non_influential_omitted)

non_influential_omitted_model
```
By removing the non-influential outlier (18), the intercept is now 1.591, compared to 1.416 previously. Percentage Change of 12.36%

```{r}
((1.591-1.416)/1.416) * 100
((1.223-1.221)/1.223) * 100
```

#The coefficient is now 1.221, compared to 1.223 previously (percentage change of 0.16%). 

```{r}
non_influential_omitted %>% 
  ggplot(aes(x = estimated_length, y = actual_length)) +
  geom_point() +
  geom_abline(
    intercept = non_influential_omitted_model$coefficients[1],
    slope = non_influential_omitted_model$coefficients[2],
    col = "red"
  ) +
  geom_abline(
    intercept = model$coefficients[1],
    slope = model$coefficients[2],
    col = "blue"
  )
```

The regression line from the data with the omitted outlier is extremely similar to the regression line for the full data set.

Obtaining the intercept and regression coefficient of estimated_length for a simple linear model fitted to data omitting one of the influential outlier points (5).
```{r}
influential_omitted <- project_management %>% 
  slice(-5)

influential_omitted_model <- lm(formula = actual_length ~ estimated_length, data = influential_omitted)

influential_omitted_model
```
By removing the non-influential outlier (5), the intercept is now 4.381, compared to 1.416 previously. Percentage Change of 209.39%

```{r}
((4.381-1.416)/1.416) * 100
((1.223-1.011)/1.223) * 100
```

#The coefficient is now 1.011, compared to 1.223 previously (percentage change of 17.33%). 

```{r}
influential_omitted %>% 
  ggplot(aes(x = estimated_length, y = actual_length)) +
  geom_point() +
  geom_abline(
    intercept = influential_omitted_model$coefficients[1],
    slope = influential_omitted_model$coefficients[2],
    col = "red"
  ) +
  geom_abline(
    intercept = model$coefficients[1],
    slope = model$coefficients[2],
    col = "blue"
  )
```
As can be seen from the plot above, by omitting the influential outlier, the regression line differs greatly to that of the regression line obtained from the full data set.
