---
title: "Text Mining"
output: html_notebook
---

```{r}
library(tidytext)
library(tidyverse)
library(janeaustenr)
```

Showing the top 20 words in Pride & Prejudice
```{r}
#generating a tibble for Pride & Prejudice book from Jane Austen package 
book_pride <- tibble(
    text = prideprejudice,
    book = "pride_prejudice",
    sentence = 1:length(prideprejudice)
  ) %>%
  unnest_tokens(word, text) 

book_pride
```

```{r}
top_20_pride <- book_pride %>% 
  #ordering by the most popular words
  count(word, sort = TRUE) %>% 
#showing the top 20 words
  slice_max(n, n = 20)


top_20_pride
```


Showing the top 20 words in Sense & Sensibility

```{r}
#generating a tibble for Sense & Sensibility book from Jane Austen package 
book_sense <- tibble(
    text = sensesensibility,
    book = "sense_sensibility",
    sentence = 1:length(sensesensibility)
  ) %>%
  unnest_tokens(word, text) 

book_sense
```

```{r}
top_20_sense <- book_sense %>% 
  #ordering by the most popular words
  count(word, sort = TRUE) %>% 
#showing the top 20 words
  slice_max(n, n = 20)

top_20_sense

```

```{r}
#Finding the most common words in Pride & Prejudice, excluding stop words
book_pride %>%
  anti_join(stop_words) %>%
  count(word) %>%
  slice_max(n, n = 20)
```


```{r}
#Finding the most common words in Sense & Sensibility, excluding stop words
book_sense %>%
  anti_join(stop_words) %>%
  count(word) %>%
  slice_max(n, n = 20)
```

Finding the most sentiment words in Pride & Prejudice

```{r}
pride_sentiment <- book_pride %>%
  inner_join(get_sentiments(lexicon = "bing")) %>%
  group_by(sentiment) %>% 
  count(word, sort = TRUE) 

pride_sentiment
```


#Generating a word cloud for Pride & Prejudice
```{r}
ggwordcloud(
  words = pride_sentiment$word,
  freq = pride_sentiment$n,
  min.freq = 20
)
```


#Adding sentiments to Sense & Sensibility
```{r}
sense_sentiment <- book_sense %>%
  inner_join(get_sentiments(lexicon = "bing")) 

sense_sentiment
```
Finding the most sentiment words in Sense & Sensibility
```{r}
sense_most_sentiment <- sense_sentiment %>%
  group_by(sentiment) %>% 
  count(word, sort = TRUE) 

sense_most_sentiment
```

#Generating a word cloud for Sense & Sensibility
```{r}
ggwordcloud(
  words = sense_most_sentiment$word,
  freq = sense_most_sentiment$n,
  min.freq = 20
)
```

#Binding both books
```{r}
books <- bind_rows(book_pride, book_sense)
```

#Joining "afinn" sentiment to books
```{r}
books_sentiments <- books %>% 
  inner_join(get_sentiments("afinn"))
```


#Creating a plot of the sentiment words both books Sense & Sensibility
```{r}
sentence_sentiments <- books_sentiments %>% 
  group_by(book) %>%
  mutate(
    word_n = row_number()
  ) %>%
  mutate(
    story_position = word_n/max(word_n) # all books on scale from 0 to 1
  ) %>% 
ggplot +
  aes(x = story_position, y = value, colour = book) +
  geom_smooth(se = FALSE) +
  guides(colour = FALSE) +
  facet_wrap(~book, nrow = 5)

sentence_sentiments
```



