---
title: "Clustering"
output: html_notebook
---

```{r}
library(tidyverse)
library(janitor)
mall_data <- read.csv("data/mall_customers.csv") %>% 
  clean_names() %>% 
  column_to_rownames("customer_id") 
```

```{r}
head(mall_data)
```

```{r}
glimpse(mall_data)
```

```{r}
#checking for missing values
summary(mall_data)

#no missing values
```

```{r}
mall_data %>%
  group_by(gender) %>%
  summarise(count = n()) %>%
ggplot(aes(x = gender, y = count)) +
  geom_col()
```

```{r}
ggplot(mall_data, aes(x = annual_income_k)) +
  geom_histogram()
```
```{r}
ggplot(mall_data, aes(x = spending_score_1_100)) +
  geom_histogram()
```
```{r}
ggplot(mall_data, aes(x = age)) +
  geom_histogram(bins = 50)
```


```{r}
#visualising the data
#exploring the trend/relationship between annual income and spending score
ggplot(mall_data, aes(x = annual_income_k, y = spending_score_1_100)) +
  geom_point()
```
There are visible clustering patterns just by eyeballing the data points in the point plot.
```{r}
#exploring the trend/relationship between age and spending score
ggplot(mall_data, aes(x = age, y = spending_score_1_100)) +
  geom_point()
```

We are interested in creating a marketing campaign to target customers based on their spending score and annual income. Perform a k-means clustering to find if there are meaningful clusters in the data to target the customers.

#1. scaling the data
```{r}
mall_scale <- mall_data %>%
                mutate_if(is.numeric, scale) %>% 
  #removing the non numerical column from the data as kmeans clustering can only work with numerical data for mathematical calulations. 
  select(-c(gender, age))

mall_scale
```
```{r}
clustered_mall <- kmeans(mall_scale, 
                            centers = 6, 
                            nstart = 25)
clustered_mall
```
6 clusters have been established.

```{r}
library(broom)

tidy(clustered_mall, 
     col.names = colnames(mall_scale))
```
```{r}
#augment gives which of the 6 clusters each observation in the original data set has been assigned to
augment(clustered_mall, mall_data)
```

```{r}
library(animation)

mall_scale %>% 
  kmeans.ani(centers = 6) 
```

```{r}
glance(clustered_mall)
```

```{r}
#Using the elbow method to choose the number of clusters
library(broom)

# Set min & max number of clusters want to look at 
max_k <- 20 

k_clusters <- tibble(k = 1:max_k) %>%
  mutate(
    kclust = map(k, ~ kmeans(mall_scale, .x, nstart = 25)), 
    tidied = map(kclust, tidy),
    glanced = map(kclust, glance),
    augmented = map(kclust, augment, mall_scale)
  )

k_clusters
```

```{r}
clusterings <- k_clusters %>%
  unnest(glanced)

clusterings
```

```{r}
ggplot(clusterings, aes(x=k, y=tot.withinss)) +
  geom_point() +
    geom_line() +
    scale_x_continuous(breaks = seq(1, 20, by = 1))
```

The plot shows that there is an elbow point at k = 5, where the rate of change of the sum of squares begins to decline. 

```{r}
library(factoextra)
```

```{r}
fviz_nbclust(mall_scale, 
             kmeans, 
             method = "wss", 
             nstart = 25)
```

I would still say the optimum number of clusters is 5.

```{r}
#Checking silhouette method
library(cluster)

#Chosen k=4 as an example 
cluster_list_k4 <-  clusterings %>% 
  unnest(augmented) %>%
  filter(k == 4) %>%
   select(.cluster) %>%
    pull()
  
 
sil <- silhouette(as.numeric(cluster_list_k4), 
                  dist(mall_scale))

fviz_silhouette(sil)
```

```{r}
fviz_nbclust(mall_scale, 
             kmeans, 
             method = "silhouette", 
             nstart = 25)
```
The silhouette method confirms my original thoughts of k = 5 being the optimum number of clusters.

```{r}
#also trying the gap start method
fviz_nbclust(mall_scale, kmeans, method = "gap_stat", nstart = 25)
```



```{r}
 clusterings %>% 
  unnest(cols = c(augmented)) %>%
  filter(k <= 5) %>%
 ggplot(aes(x = annual_income_k, y = spending_score_1_100)) +
  geom_point(aes(color = .cluster)) + 
  facet_wrap(~ k)
```
k = 5 definitely represents the clustering of the data points efficiently.

```{r}
#Looking at k = 5 specifically
 clusterings %>% 
  unnest(cols = c(augmented)) %>%
  filter(k == 5) %>%
 ggplot(aes(x = annual_income_k, y = spending_score_1_100, colour = .cluster, label = age)) +
  geom_point(aes(color = .cluster)) +
  geom_text(hjust = 0, vjust = - 0.5, size = 3)
```

```{r}
 clusterings %>% 
  unnest(augmented) %>%
  filter(k == 5) %>%
  group_by(.cluster) %>%
  summarise(mean(annual_income_k), mean(spending_score_1_100))
```


Comment on the attributes on one or two of the clusters (maybe even give them a label if you like - like in section 4.1 of the ‘Segmentation & clustering intro’ lesson).

Cluster number 5 is high income and high spend - these may be more 'high end' customers and could look to target them accordingly. 
#Cluster number 2 is high income but low spend - perhaps they mainly shop at other malls or spend money on activities/items not sold within your mall. May want to target with further data collection to find out more about this in case we are able to target with a particular campaign if we find they are shopping at competitors. 