---
title: "Multiple Regression"
output: html_notebook
---

```{r}
library(tidyverse)
```

```{r}
diamonds <- read_csv("data/diamonds.csv")
```

```{r}
head(diamonds)
```

```{r}
glimpse(diamonds)
```

```{r}
names(diamonds)
```

```{r}
alias(lm(carat ~ ., data = diamonds))

#No alias are reported.
```

#Using ggpairs to investigate if there is a strong correlation between the physical dimensions x, y and z with carat.

```{r}
library(GGally)

ggpairs(diamonds[,c("carat", "x", "y", "z")])
```

Examining the pairs plots above, there is a very strong correlation between carat and x, y & z (0.975, 0.952 & 0.953 respectively).

```{r}
#Removing columns x, y & z from the dataset
diamonds_trim <- diamonds %>% 
  select(-c("x", "y", "z"))
```

```{r}
ggpairs(diamonds_trim)
```



Building a regression model for the price of a diamond in terms of the possible predictor variables in the dataset.
* The response will be price.
* The predictor will be carat.
* The simple linear regression equation is price^ = intercept + b(carat) x carat
* The hypothesis is that price is more likely to be higher as carat increases.


```{r}
diamonds_trim %>% 
  ggplot(aes(x = carat, y = price)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)

```

From the plot, it can be seen that there is a linear relationship: as the carat increases, the price of diamonds increases.

Checking if cut helps to predict the price of diamonds.

```{r}
diamonds_trim %>% 
  ggplot(aes(x = cut, y = price)) +
  geom_boxplot()
```
Checking if clarity helps to predict the price of diamonds.

```{r}
diamonds_trim %>% 
  ggplot(aes(x = clarity, y = price)) +
  geom_boxplot()
```

Checking if colour helps to predict the price of diamonds.

```{r}
diamonds_trim %>% 
  ggplot(aes(x = color, y = price)) +
  geom_boxplot()
```

```{r}
#Using dummy_cols() to generate dummies for cut, clarity and colour.
library(fastDummies)
library(janitor)
```

```{r}
cor.test(diamonds$depth, diamonds$carat)
```


```{r}
diamonds_dummies <- diamonds_trim %>% 
  dummy_cols(select_columns = c("cut", "color", "clarity"), remove_first_dummy = TRUE) %>% 
  clean_names()
  
```

```{r}
#Checking the number of unique values for cut, color and clarity columns
diamonds_dummies %>% 
distinct(cut)

diamonds_dummies %>% 
distinct(color)

diamonds_dummies %>% 
distinct(clarity)

#cut - 5
#colour - 7
#clarity - 8

#Comparing the number of unique values we see that when using the dummy columns function, this has taken 1 name away from each of the columns as this can be predicted by the others remaining in each column. 

#cut - 4
#colour - 6
#clarity - 7

```


```{r}
#Goodness-of-fit metrics:-

library(ggfortify)

model <- lm(price ~ carat, data = diamonds_trim)

autoplot(model)
```

```{r}
summary(model)
```

#Adding clarity to the model
```{r}
model2 <- lm(price ~ carat + clarity, data = diamonds_trim)
```

```{r}
autoplot(model2)
```

```{r}
summary(model2)
```
