---
title: "The Titanic & Probability"
output: html_notebook
---

```{r}
#Loading prob library first to ensure the dplyr filter can be recognized later 
library(prob)
```

```{r}
library(gtools)
library(tidyverse)
```

1. The Titanic

```{r}
#Loading the data
titanic_data <- read_csv("titanic.csv")

#Viewing class names
class_names <- unique(titanic_data$class)

class_names
```

```{r}
#Creating a column for crew 
titanic_data <- titanic_data %>% 
  #Ensuring 1st, 2nd & 3rd class passengers are represented as false in the crew column
  mutate(crew = case_when(class == "1st" ~ FALSE,
                          class == "2nd" ~ FALSE,
                          class == "3rd" ~ FALSE,
                          #else true
                          TRUE ~ TRUE))


```

1. Calculate the conditional probability 
prob( survived ∣ crew )

```{r}
prob_crew_survived <- titanic_data %>% 
  filter(survived == "yes", crew == TRUE) %>% 
  summarise(prob = n() / nrow(titanic_data))
prob_crew_survived
```

Probability that a person was a crew member
```{r}
crew_probability <- titanic_data %>% 
  filter(crew == TRUE) %>% 
  summarise(prob = n()) / nrow(titanic_data)
crew_probability
```

```{r}
prob_crew_survived/crew_probability
```



2. Calculate the conditional probability 
prob( survived ∣ passenger )
```{r}
prob_passenger_survived <- titanic_data %>% 
  filter(survived == "yes", crew == FALSE) %>% 
  summarise(prob = n() / nrow(titanic_data))
prob_passenger_survived
```


Probability that a person was a passenger
```{r}
passenger_probability <- titanic_data %>% 
  filter(crew == FALSE) %>% 
  summarise(prob = n()) / nrow(titanic_data)
passenger_probability
```


```{r}
prob_passenger_survived/passenger_probability
```


2. MVP and Extension - Probability problems
1. If I roll a six-sided dice once, what is the probability of getting a number less than 5?
```{r}
rolldie(1) %>% 
  filter(X1 < 5)


#There are 6 total outcomes and 4 opportunities to roll a number lower than 5.
#The Probability is 4/6
4/6
```

2. How many outcomes are there in the sample space of an experiment in which we toss three dice?
```{r}
rolldie(3) %>% 
  nrow()

#There are 216 possible outcomes of rolling 3 die.
```

3. In an experiment where we roll four dice:

```{r}
rolldie(4) %>% 
  nrow()
```


i. how many outcomes are in the event total_equals_5_6_7 (i.e. the total of the four dice equals 5, 6 or 7)?
```{r}
total_equals_5_6_7 <- 
rolldie(4) %>% 
  filter(X1 + X2 + X3 + X4 == 5 | X1 + X2 + X3 + X4 == 6 | X1 + X2 + X3 + X4 == 7) 
#There are 34 outcomes where the total of four die equals 5, 6 or 7
```
`


ii. what is the probability of this event?
```{r}
#The probability of this event is 
34/1296
```


iii. what assumptions do we apply in answering part ii?
- Assuming that the order that the dice are rolled does not matter

4. Five chairs are placed in a row in a room, along with five people of differing height. The people sit down randomly on the chairs. What is the probability that they sit in order of decreasing height, left to right?

```{r}
#Outcomes where height order matters
five_chairs_permutations <- as_tibble(
permutations(
  n = 5,
  r = 5
)
) 
#There are 120 possible outcomes of five people sitting in a random order

five_chairs_combinations <- as_tibble(
combinations(
  n = 5,
  r = 5,
  repeats.allowed = FALSE
)
) 
#There is one possible outcome of people sitting in decreasing height order

#The probability is the possibility of the specified outcome divided by the total number of all possible outcomes (1/120)
nrow(five_chairs_combinations)/nrow(five_chairs_permutations)

```

5. I roll six dice:

```{r}
rolldie(6) %>% 
  nrow
#There are 6 dice (n = 6) and we are choosing 6 at a time (r = 6)
six_dice <- 
rolldie(6) %>% 
  as_tibble(permutations(
    n = 6,
    r = 6,
    repeats.allowed = TRUE
  )
  ) 
```


1. what is the probability I see the numbers 6, 5, 4, 3, 2 and 1 in order on the dice, left to right?
```{r}
ordered_six <- six_dice %>% 
  isin(., c(6, 5, 4, 3, 2, 1), ordered = TRUE) 

sum(ordered_six)/nrow(six_dice)
  
```


2. what is the probability I see one each of the numbers 6, 5, 4, 3, 2 and 1 on the dice regardless of order?

```{r}
random_order <- six_dice %>% 
  isin(., c(6, 5, 4, 3, 2, 1), ordered = FALSE) 

sum(random_order)/nrow(six_dice)
```

