---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(janitor)
qb_revenue_breakdown <- CodeClanData::qb_revenue_breakdown %>% clean_names
qb_monthly_sales <- CodeClanData::qb_monthly_sales %>% clean_names()
qb_competitors <- CodeClanData::qb_competitors %>% clean_names()
qb_device_data <- CodeClanData::qb_device_data %>% clean_names()
```

4.1 Part 1
1. Take the data in the dataset qb_revenue_breakdown and make a stacked bar chart showing the sources of revenue across the two years in the dataset.
```{r}
ggplot(qb_revenue_breakdown) +
  geom_col(aes(x = year, y = revenue, fill = product)) +
  theme_light() +
  scale_y_continuous(labels = scales::comma) +
  scale_fill_manual(values = c("#E89FE9", "#50434F", "#B6A7B5", "#F9A472", "#BD7040")) +
  labs(x = "Year",
       y = "Revenue",
       title = "Sources of Revenue in 2018 & 2019") +
  theme(title = element_text(size = 13)) +
  theme(text = element_text(size = 12)) +
  theme(axis.text = element_text(size = 11))
```

2. Make a line chart showing monthly sales of the “flagship product” - the ‘KwikBit Unit 1’ - over the last year. This data is inside qb_monthly_sales, but you will need to use filter() or subset() (base R).
```{r}
flagship_product <- qb_monthly_sales %>% 
  filter(cashflow == "Kwikbit Unit 1 Sales") 

ggplot(flagship_product,
       aes(x = date, y = amount)) +
  geom_line() +
  labs(
    x = "Date",
    y = "Monthly Sales",
    title = "Monthly Sales of Kwikbit Unit 1"
  ) +
  theme_light() +
  scale_y_continuous(labels = scales::comma) +
  theme(title = element_text(size = 13)) +
  theme(text = element_text(size = 12)) +
  theme(axis.text = element_text(size = 11))
```
3. Make a line chart showing monthly revenue and costs over the last year. This data is also in qb_monthly_sales.
```{r}

revenue_costs <- qb_monthly_sales %>% 
  filter(cashflow %in% c("Cost", "Total Revenue"))

ggplot(revenue_costs) +
 geom_line(aes(x = date, y = amount, colour = cashflow)) +
  theme_light() +
  scale_y_continuous(labels = scales::comma) +
  scale_y_continuous(limits = c(0, 3000)) +
  theme(title = element_text(size = 13)) +
  theme(text = element_text(size = 12)) +
  theme(axis.text = element_text(size = 11)) +
  scale_colour_manual(values = c("#E89FE9", "#50434F", "#B6A7B5", "#F9A472", "#BD7040"))
```

4. Show annual sales of personal fitness trackers over the last 5 years broken down by company as a ribbon plot (use geom_area). This data is in qb_competitors.

```{r}
ggplot(qb_competitors) +
  geom_area(aes(x = year, y = revenue, fill = company)) +
  labs(
    title = "Competitor Sales over Time"
  ) +
  theme_light() +
  scale_y_continuous(labels = scales::comma) +
  scale_fill_manual(values = c("#E89FE9", "#50434F", "#B6A7B5", "#F9A472", "#BD7040")) +
  theme(title = element_text(size = 13)) +
  theme(text = element_text(size = 12)) +
  theme(axis.text = element_text(size = 11))
```

5. Now show the sales from the four competitors as a line graph. Include an extra layer that shows the data points used to make the lines.
```{r}
ggplot(qb_competitors) +
  aes(x = year, y = revenue, colour = company) +
  geom_line() +
  geom_point() +
  labs(
    title = "Competitor Sales over Time"
  ) +
  theme_light() +
  scale_y_continuous(labels = scales::comma) +
  scale_colour_manual(values = c("#E89FE9", "#50434F", "#B6A7B5", "#F9A472", "#BD7040")) +
  theme(title = element_text(size = 13)) +
  theme(text = element_text(size = 12)) +
  theme(axis.text = element_text(size = 11))
```
6. Now the company wants to compare the number of steps that their device counts vs. their competitors. Make a line graph of the number of steps throughout time, and use faceting to compare between companies and people. The data for this is in qb_device_data.
```{r}
qb_device_data <- qb_device_data %>% 
  mutate(time = hours * 60 + mins)

ggplot(qb_device_data) +
  aes(x = time, y = counts, colour = device) +
  geom_line() +
  facet_grid(~ id ~ device) +
  labs(
    x = "Time (mins)",
    y = "Number of Steps",
    title = "Number of Steps throughout Time",
    subtitle = "Comparison between Qikbit1 & Competitors") +
  theme_light() +
  scale_colour_manual(values = c("#E89FE9", "#50434F", "#B6A7B5", "#F9A472", "#BD7040")) +
  theme(title = element_text(size = 13)) +
  theme(text = element_text(size = 12)) +
  theme(axis.text = element_text(size = 11))
```

