---
title: "Manual Model Development"
output: html_notebook
---

#Building an explanatory model of housing sale prices for the last few years in King Country (near Seattle) between May 2014 and May 2015.

```{r}
#Loading library and reading in the data
library(tidyverse)
houses <- read_csv("data/kc_house_data.csv")
```

```{r}
names(houses)
```

```{r}
glimpse(houses)
```

```{r}
#Removing irrelevant columns
houses_clean <- houses %>% 
  select(-c("id", "sqft_living15", "sqft_lot15", "zipcode"))
```

```{r}
#Amending "waterfront" column to show True or False values (1 or 0)
houses_clean <- houses_clean %>% 
  mutate(waterfront = as.logical(waterfront))
```

```{r}
#Creating a new renovated column to show True for renovated and False if property has not been renovated and replacing yr_renovated column with this.
houses_clean <- houses_clean %>% 
 mutate(renovated = yr_renovated != 0) %>%
  select(-"yr_renovated") 
```


```{r}
#Condition and grade columns are categorical ordinal data types, not interval. 
#grade - An index from 1 to 13, where 1-3 falls short of building construction and design, 7 has an average level of construction and design, and 11-13 have a high quality level of construction and design.
#condition - An index from 1 to 5 on the condition of the apartment.
  houses_clean <- houses_clean %>%
  mutate(condition = as_factor(condition)) %>%
  mutate(grade = as_factor(grade))
```



```{r}
#Checking for aliased variables (., - with all other variables)

alias(price ~ ., houses_clean)

#It can be seen that sqft_basement can be computed as:-
#sqft_basement = no sqft_above + sqft_living

#sqft_living is an aliased variable.
```
```{r}
#Dropping aliased variable sqft_living
houses_clean <- houses_clean %>% 
  select(-sqft_living)
```


```{r}
#Checking if all aliased variables have been removed
alias(price ~ ., houses_clean)

#All aliased variables have been removed.
```


```{r}
#Splitting the dataset into numeric and non-numeric columns to assist ggpairs() to run in manageable time.
#Adding price column to the non-numeric dataframe in order to see it's correlation with the non-numeric predictors.

library(GGally)
houses_tidy_numeric <- houses_clean %>%
  select_if(is.numeric)

houses_tidy_nonnumeric <- houses_clean %>%
  select_if(function(x) !is.numeric(x))

houses_tidy_nonnumeric$price <- houses_clean$price

ggpairs(houses_tidy_numeric)
ggpairs(houses_tidy_nonnumeric)
```

The plot above shows that sqft_above is strongly correlated with price. Grade and waterfront with price looks good as som separation can be seen.

```{r}
houses_clean %>%
  ggplot(aes(x = grade, y = price)) +
  geom_boxplot()
```

```{r}
houses_clean %>%
  ggplot(aes(x = waterfront, y = price)) +
  geom_boxplot()
```


```{r}
library(modelr)
```

```{r}
#Building a regression model for the price response variable in terms of the multiple potential predictors held in the data frame.

#First predictor is sqft_living
mod1a <- lm(price ~ sqft_above, data = houses_clean)
summary(mod1a)
```

```{r}
mod1b <- lm(price ~ grade, data = houses_clean)
summary(mod1b)
```


```{r}
mod1c <- lm(price ~ waterfront, data = houses_clean)
summary(mod1c)
```
```{r}
# grade looks the most promising, but some of the grade level coefficients are insignificant.
# the F-test at the bottom of the regression output tests against the null model (i.e. intercept only)
# but, if we want, we can replicate this using a separate anova
# null model: regress price on intercept only
null_model <- lm(price ~ 1, data = houses_clean)
grade_model <- lm(price ~ grade, data = houses_clean)
anova(null_model, grade_model)

```

```{r}
# grade is significant, let's keep it. Now plot diagnostics
par(mfrow = c(2, 2))
plot(mod1b)
```

```{r}
houses_resid <- houses_clean %>%
  add_residuals(mod1b) %>%
  select(-c("price", "grade"))

houses_resid_numeric <- houses_resid %>%
  select_if(is.numeric)

houses_resid_nonnumeric <- houses_resid %>%
  select_if(function(x) !is.numeric(x))

houses_resid_nonnumeric$resid <- houses_resid$resid
```


```{r}
ggpairs(houses_resid_numeric)

```


```{r}
ggpairs(houses_resid_nonnumeric)

```

lat has highest correlation with residuals, but, again, waterfront still looks pretty promising.

```{r}
#to view specific columns
ggpairs(houses_tidy_numeric, columns = c(1, 6, 2))
```


```{r}
mod2a <- lm(price ~ grade + lat, data = houses_tidy)
summary(mod2a)
```
```{r}
# lat is significant and higher r^2, let's keep model2a
par(mfrow = c(2, 2))
plot(mod2a)
```


```{r}
#Checking remaining residuals
houses_resid <- houses_clean %>%
  add_residuals(mod2a) %>%
  select(-c("price", "grade", "lat"))

houses_resid_numeric <- houses_resid %>%
  select_if(is.numeric)

houses_resid_nonnumeric <- houses_resid %>%
  select_if(function(x) !is.numeric(x))

houses_resid_nonnumeric$resid <- houses_resid$resid
```

```{r}
ggpairs(houses_resid_numeric)

```

```{r}
ggpairs(houses_resid_nonnumeric)
```

```{r}
mod3a <- lm(price ~ grade + lat + view, data = houses_clean)
summary(mod3a)
```


```{r}
mod3b <- lm(price ~ grade + lat + waterfront, data = houses_clean)
summary(mod3b)

```


```{r}
# view model is best, keep mod3a
par(mfrow = c(2, 2))
plot(mod3a)
```

```{r}
houses_resid <- houses_clean %>%
  add_residuals(mod3a) %>%
  select(-c("price", "grade", "lat", "view"))

houses_resid_numeric <- houses_resid %>%
  select_if(is.numeric)

houses_resid_nonnumeric <- houses_resid %>%
  select_if(function(x) !is.numeric(x))

houses_resid_nonnumeric$resid <- houses_resid$resid
```
```{r}
ggpairs(houses_resid_numeric)

```


```{r}
ggpairs(houses_resid_nonnumeric)

```

sqft_basement has highest correlation with residuals. Testing against all remaining categorical predictors:

```{r}
mod4a <- lm(price ~ grade + lat + view + sqft_basement, data = houses_clean)
summary(mod4a)
```

```{r}
mod4b <- lm(price ~ grade + lat + view + waterfront, data = houses_clean)
summary(mod4b)
```
```{r}
mod4c <- lm(price ~ grade + lat + view + condition, data = houses_clean)
summary(mod4c)
```



```{r}
mod4d <- lm(price ~ grade + lat + view + renovated, data = houses_clean)
summary(mod4d)
```

```{r}
# looks like model with sqft_basement is best, keep mod4a
par(mfrow = c(2, 2))
plot(mod4a)
```

```{r}
houses_resid <- houses_clean %>%
  add_residuals(mod4a) %>%
  select(- price)
```
price ~ grade + lat + view + sqft_basement is the best model

```{r}
mod5a <- lm(price ~ grade + lat + view + sqft_basement + grade:lat, data = houses_clean)
summary(mod5a)
```



```{r}
mod5b <- lm(price ~ grade + lat + view + sqft_basement + grade:view, data = houses_clean)
summary(mod5b)
```

```{r}
mod5c <- lm(price ~ grade + lat + view + sqft_basement + grade:sqft_basement, data = houses_clean)
summary(mod5c)
```

```{r}
mod5d <- lm(price ~ grade + lat + view + sqft_basement + lat:view, data = houses_clean)
summary(mod5d)
```

```{r}
mod5e <- lm(price ~ grade + lat + view + sqft_basement + lat:sqft_basement, data = houses_clean)
summary(mod5e)
```

```{r}
mod5f <- lm(price ~ grade + lat + view + sqft_basement + view:sqft_basement, data = houses_clean)
summary(mod5f)
```

```{r}
# mod5c looks like the best
par(mfrow = c(2,2))
plot(mod5c)
```

```{r}
houses_resid %>%
  ggplot(aes(x = sqft_basement, y = resid, colour = grade)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE) +
  facet_wrap(~ grade)
```

